---
title: "Hcrt analysis - males only"
output: pdf_document
---

```{r opts, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "male_images_gr/"
)
```


```{r}
library(Seurat)
library(ggplot2)
library(sctransform)
library(cowplot)
library(dplyr)
library(scrattch.hicat)

```

load only Hcrt core cells, with doublets removed
```{r}
load('~/postdoc2/Shibin_hypocretin/hypocretin/hcrtcorecells.RData')
```
relabel clusters
```{r}
hcrt<-RenameIdents(hcrt,"4"="1")
hcrt<-RenameIdents(hcrt,"10"="2")
hcrt<-RenameIdents(hcrt,"9"="3")
hcrt$merged.res.2.renamed<-Idents(hcrt)
hcrt$merged.res.2.renamed<-factor(hcrt$merged.res.2.renamed,levels=c("0","1","2","3"))
Idents(hcrt)<-'merged.res.2.renamed'
```
split by sex - males here:
```{r}
sex<-as.factor(hcrt$librarynames %in% c("YF2","OF2"))
levels(sex)<-c('male','female')
hcrt$sex<-sex
Idents(hcrt)<-'sex'
hcrt<-subset(hcrt,idents='male')

```



check genes and umis detected per library -- data quality

```{r}
VlnPlot(hcrt,features=c("nFeature_RNA","nCount_RNA"),group.by='librarynames',pt.size = 0,cols=c("gray40","red"))
```


```{r,fig.width=15,fig.height=6}
p1<-DimPlot(hcrt, label = TRUE,group.by='merged.res.2.renamed') + NoLegend()
p2<-DimPlot(hcrt, label = TRUE,group.by='age',cols=c("gray40","red"))
p3<-DimPlot(hcrt, label = FALSE,group.by='librarynames',cols=c("gray40","red")) 
plot_grid(p1,p2,p3,ncol=3)
```
check genes and umis detected per cluster -- data quality + some notion of how large a cell we are looking at.

```{r}
VlnPlot(hcrt,features=c("nFeature_RNA","nCount_RNA"),group.by='merged.res.2.renamed',pt.size = 0)
```
Look for cluster specific genes:
```{r}
Idents(hcrt)<-'merged.res.2.renamed'
markers<-FindAllMarkers(hcrt,logfc.threshold = log(2))
markers %>% group_by(cluster) %>% top_n(5,avg_logFC)
```
```{r}
#markers.filtered<-markers[markers$p_val_adj<0.05,]
#write.csv(markers.filtered,file='~/postdoc2/Shibin_hypocretin/hypocretin/males_cluster_DEGs.csv')
```


plot a heatmap
```{r}
markers %>% group_by(cluster) %>% top_n(10,avg_logFC) ->top10
DoHeatmap(hcrt, features = top10$gene,slot='data')+
  scale_fill_gradientn(colors = rev(RColorBrewer::brewer.pal(n = 10, name = "Greys")))
```
```{r}
DoHeatmap(hcrt, features = top10$gene,slot='data')+
  scale_fill_gradientn(colors = rev(RColorBrewer::brewer.pal(n = 10, name = "Oranges")))
```
Featureplots of marker genes

```{r,fig.width=15,fig.height=10}
FeaturePlot(hcrt,features=c("Snap25",'Hcrt',"Hmcn1","Cntn4","Qrfp","Sgcd","Adgrb3","Tmem108","Col25a1"))
```
```{r,fig.width=15,fig.height=10}
VlnPlot(hcrt,features=c("Snap25",'Hcrt',"Hmcn1","Cntn4","Qrfp","Sgcd","Adgrb3","Tmem108"),group.by = "merged.res.2.renamed",ncol=3)
```

Mickelsen's genes:
```{r,fig.width=15,fig.height=10}
FeaturePlot(hcrt,c("Hcrt", "Pdyn", "Npvf", "Rfx4", "Scg2", "Nptx2", "Lhx9", "Pcsk1", "Nek7", "Plagl1", "Vgf", "Slc2a13"))
```

do young and old cells contribute differently to the 4 clusters?
total cell numbers.
```{r}
table(hcrt$merged.res.2.renamed,hcrt$librarynames)
```
percentage.
```{r}
t<-table(hcrt$merged.res.2.renamed,hcrt$librarynames)
prop.table(t,2) # cell percentages

```
the distribution of cells across the 4 clusters from each library is remarkably similar.



Lets look for differentially expressed genes between young and old in the supercluster of 0,1,2


```{r}
hcrtcore<-subset(hcrt,idents=c("0","1","2"))
```

```{r}
table(hcrtcore$age)
```
```{r}
table(hcrtcore$librarynames)
```

first though, Mickelsen's genes:
```{r,fig.width=15,fig.height=10}
FeaturePlot(hcrtcore,c("Hcrt", "Pdyn", "Npvf", "Rfx4", "Scg2", "Nptx2", "Lhx9", "Pcsk1", "Nek7", "Plagl1", "Vgf", "Slc2a13"))
```
```{r,fig.width=15,fig.height=10}
Idents(hcrtcore)<-'age'
VlnPlot(hcrtcore,c("Hcrt", "Pdyn", "Npvf", "Rfx4", "Scg2", "Nptx2", "Lhx9", "Pcsk1", "Nek7", "Plagl1", "Vgf", "Slc2a13"),cols=c("gray40","red"))
```

New plots for Shibin. 
```{r,fig.width=15,fig.height=10}
genesofinterest<-c("Hcrt","Scg2","Penk","Gal","Jak1","Nek7","Nptx2","Slc2a13","Nucb2","Slc17a6","Pdyn")
VlnPlot(hcrtcore,genesofinterest,idents = 'young',y.max=7,pt.size = 0,cols=c("gray40"))
```
New plots for Shibin. 
```{r,fig.width=15,fig.height=10}
VlnPlot(hcrtcore,genesofinterest,idents = 'old',y.max=7,pt.size=0,cols=c("red"))
```




New plots of KCNQ channels 
```{r,fig.width=15,fig.height=10}
genesofinterest<-c("Kcnq1","Kcnq2","Kcnq3","Kcnq4","Kcnq5")
VlnPlot(hcrtcore,genesofinterest,y.max=4,pt.size = 0,cols=c("gray40","red"))
```

Fraction expressed.

```{r}
a <- DotPlot(object = hcrtcore, features = genesofinterest)
a$data[,c('features.plot','id','pct.exp')]
```

manually double check this:

```{r}
my_cells<-WhichCells(hcrtcore,idents = 'young')
sum(GetAssayData(object = hcrtcore, slot = "data")['Kcnq1',my_cells]>0)/length(my_cells)
```
ok all good.
```{r}
Idents(hcrtcore)<-'age'
agemarkers<-FindMarkers(hcrtcore,ident.2="young",ident.1="old")
#agemarkers
```
plot heatmap of sorted genes.
```{r}
a<-agemarkers[agemarkers$p_val_adj<0.05,]
b<-agemarkers[order(a$avg_logFC,decreasing=T),]
DoHeatmap(hcrtcore,features=rownames(b),group.by='librarynames',slot='data')+
  scale_fill_gradientn(colors = rev(RColorBrewer::brewer.pal(n = 9, name = "Greys")))
```
```{r}
DoHeatmap(hcrtcore,features=rownames(b),group.by='librarynames',slot='data')+
  scale_fill_gradientn(colors = rev(RColorBrewer::brewer.pal(n = 9, name = "Oranges")))
```


plot all significant makers.
```{r,fig.width=15,fig.height=6}
Idents(hcrtcore)<-"age"
VlnPlot(hcrtcore,features=rownames(b),pt.size = 0,cols=c("gray40","red"))
```


what does this look like in a volcano plot?
```{r}
#try DEseq DEG test
agemarkers2<-FindMarkers(hcrtcore,ident.2="young",ident.1="old",logfc.threshold = 0)
#agemarkers2
```
```{r,fig.height=6,fig.width=6}
# convert from ln to log2 fold change.
agemarkers2$avg_log2FC<-agemarkers2$avg_logFC*log2(exp(1))
library(EnhancedVolcano)
EnhancedVolcano(agemarkers2,
                lab=rownames(agemarkers2),
                x='avg_log2FC',
                y='p_val_adj',
                xlim=c(-1,2.3),
                FCcutoff=0.25,
                transcriptPointSize=3,
                transcriptLabSize = 5)
```

```{r}
write.csv(agemarkers2[agemarkers2$p_val_adj<0.05,],file = '~/postdoc2/Shibin_hypocretin/hypocretin/agemarkers_male.csv')
```






